Automated alpha mining has recently shifted from operator- and formula-based search toward large language model (LLM) agents. However, many LLM-based systems lack an explicitly defined and learnable hypothesis space, relying on agents to generate,memorize, reflect, and propose new search directions. This entangles semantic exploration with code generation and makes search feedback difficult to learn and reuse. We introduce \method{}, which reformulates alpha mining as exploration over an explicit space of trading semantics. Each hypothesis is represented by a structured schema comprising \emph{Event}, \emph{Context}, \emph{Qualities}, \emph{Direction}, and \emph{Output}. \method{} decouples exploration from realization: an LLM translates selected schemas into leakage-checked executable factor code, while exploration is learned from evaluated schema--reward pairs and performed by a transparent selector over promising, uncertain, novel, and locally mutated hypotheses. Experiments on Chinese stock market show that \method{} discovers factor pools with strong predictive and portfolio performance. Our analyses show that the trading-semantic space is not only learnable, but also meaningfully navigable: by balancing exploration and exploitation, the search process identifies high-value semantic regions without prematurely collapsing onto known patterns. Moreover, within our framework, the quality of successfully generated factors is largely decoupled from the capability of the underlying LLM. This enables scalable alpha mining with substantially more economical models.

Automated alpha mining has recently shifted from operator- and formula-based search toward large language model (LLM) agents. However, many LLM-based systems lack an explicit exploration space and a principled search policy, relying instead on LLM agents to infer new directions through generation and self-reflection. As a result, exploration remains largely implicit and difficult to control or optimize systematically. We introduce \method{}, which constructs and explores a structured space of trading semantics for alpha mining. Each point in this space is a schema plan composed of \emph{Event}, \emph{Context}, \emph{Qualities}, \emph{Direction}, and \emph{Output}, specifying the semantics of a candidate factor before implementation. \method{} decouples exploration from realization: an LLM translates selected schemas plan into leakage-checked executable factors, while a learned selector explores evaluated schema--reward relationships across promising, uncertain, novel, and locally mutated hypotheses. Experiments on the Chinese stock market show that \method{} discovers factor pools with strong predictive and portfolio performance. Further analyses show that the semantic space is learnable and can be navigated by balancing exploration and exploitation toward high-value regions. Within our framework, factor quality is largely decoupled from LLM capability, enabling scalable alpha mining with more economical models.