\section{Introduction}
\label{sec:introduction}

Alpha mining aims to discover trading signals that are predictive, executable, interpretable, and stable out of sample. Traditional asset-pricing research derives factors from economic hypotheses and
empirical anomalies \cite{sharpe1964capital,fama1992cross,fama1993common},but the vast and evolving space of
potential trading mechanisms makes manual discovery inherently limited.\textcolor{red}{[citation needed]}. An automated factor-mining system should therefore not only generate candidate signals, but also efficiently explore meaningful trading hypotheses while preserving economic interpretation, implementation validity, and reproducible evaluation.

Existing automated systems have expanded alpha discovery through formulaic and evolutionary search. Formulaic and evolutionary methods represent factors as expressions, syntax trees, or operator programs \cite{kakushadze2016formulaic,koza1992genetic,su2022genetic,zhang2020autoalpha,cui2021alphaevolve,yu2023alphagen,shi2025alphaforge}, enabling large-scale exploration of implementation spaces. However, the search process is primarily driven by mathematical composition rather than explicit economic semantics, making it difficult to
incorporate higher-level trading hypotheses into the discovery process. 

More recent LLM-based agents introduce planning, code generation, memory system, trajectory evolution, and guarded execution into factor discovery \cite{li2024can,tang2025alphaagent,shi2025navigating,lin2026factorengine,han2026quantaalpha,shi2026hubble,liu2025cogalpha,chen2025alphasage}. However, existing LLM-based systems often delegate both factor realization and exploration decisions to the LLM agent itself. While agentic workflows introduce semantic reasoning into factor discovery, they typically lack an explicitly defined search space, principled exploration algorithms, and structural constraints on factor construction. As a result, the discovery process becomes intertwined with model-specific reasoning traces and workflow designs, making exploration difficult to learn, diagnose, and reproduce. Moreover, the effectiveness of such systems is closely tied to the capability of the underlying LLMs, requiring expensive frontier models and limiting the scalability of large-scale automated mining.

We propose \method{}, a semantic schema-guided evolution framework that separates \emph{what to search} from \emph{how to implement it}. Instead of treating code or formulas as the primary search object, \method{} represents each candidate as a structured trading-semantic plan
\begin{equation}
p=(e,c,Q,d,o),
\end{equation}
where \(e\) denotes a market event, \(c\) its conditioning context, \(Q\) a set of quality constraints, \(d\) the directional hypothesis, and \(o\) the output form. The search process operates over this schema space, while an LLM translates selected plans into executable factor implementations for evaluation.

\method{} maintains a reward buffer of evaluated plans and trains a LightGBM ensemble surrogate over schema features \cite{ke2017lightgbm}. For unseen plans, the surrogate estimates both predicted reward and ensemble-disagreement uncertainty. A quota selector allocates the limited implementation and backtesting budget across high-reward candidates, uncertain candidates, structurally novel candidates, and mutation-based refinements around historically strong plans. By associating evaluation feedback with schema plans rather than generated code or operators, \method{} learns reusable priors over which trading semantics are worth exploring.

Our contributions are:
\begin{itemize}[leftmargin=12pt]
    \item We introduce structured trading-semantic plans as a new search abstraction for alpha mining, moving exploration from implementation artifacts to explicit semantic hypotheses.
    \item We formulate semantic exploration as a learnable optimization problem, using accumulated plan evaluations and surrogate-based search to balance exploitation, uncertainty, novelty, and refinement under limited budgets.
    \item We demonstrate that the trading-semantic space is informative and navigable, and that factor quality is largely independent of the choice of LLM backend, enabling scalable alpha mining with economical models.
\end{itemize}