\section{Preliminary}
\label{sec:preliminary}

\subsection{Alpha Mining}

An alpha factor is a rule that transforms historical market information
into a cross-sectional signal for asset ranking or prediction. Given
\(N\) tradable assets observed over \(T\) trading dates, let
\(X_t\in\mathbb{R}^{N\times D}\) denote the market state at date \(t\),
where rows correspond to assets and columns correspond to observable
market features. A factor \(f\) maps a historical window of market
states into a signal vector:

\[
v_t=f(X_{t-\tau+1:t}),\quad v_t\in\mathbb{R}^{N}.
\]

Beyond its executable implementation, a factor typically reflects an
underlying market mechanism that provides an interpretable rationale
for its predictive relationship with future returns. Therefore, alpha
mining involves not only finding predictive transformations, but also
discovering meaningful mechanisms and realizing them as executable
factors.

Formally, we distinguish a semantic factor plan from its executable
realization. Let \(p\in\mathcal{P}\) denote a candidate trading-semantic
plan, and let

\[
R:\mathcal{P}\rightarrow\Delta(\mathcal{F})
\]

denote a realization process that maps a semantic plan to a distribution
over executable factors. A realized factor \(f\sim R(p)\) is evaluated by
a factor-quality objective \(\mathcal{J}(f)\). The objective of semantic
factor discovery can therefore be written as

\[
p^*=\arg\max_{p\in\mathcal{P}}
\mathbb{E}_{f\sim R(p)}[\mathcal{J}(f)].
\]

A plan does not uniquely determine an executable factor: different
realizations may adopt different operators, parameters, and functional
forms. Thus, the objective concerns the reward distribution induced by
a plan rather than a deterministic mapping.

\subsection{Structured Semantic Plans}

To search over trading semantics, \method{} represents each candidate
factor as a structured semantic plan. Rather than assuming a unique
ontology of market mechanisms, we define a practical and expressive
representation that captures recurring dimensions of factor
construction. A semantic plan describes the intended trading mechanism
before implementation and can subsequently be translated into an
executable factor by an LLM.

Formally, a semantic plan is represented as

\[
p=(e,c,Q,d,o)\in\mathcal{P},
\]

where the five dimensions jointly characterize different aspects of a
candidate trading mechanism:

\begin{itemize}[leftmargin=12pt, nosep]
  \item \textbf{Event}: the market phenomenon or signal-generating event
        being captured, such as breakout, volume expansion, or volatility
        compression;

  \item \textbf{Context}: the market state or reference condition under
        which the event is interpreted, such as a recent extreme, VWAP
        region, volatility regime, or cross-sectional quantile;

  \item \textbf{Qualities}: additional properties or validation criteria
        that refine the mechanism, such as volume confirmation,
        multi-horizon consistency, or outlier filtering;

  \item \textbf{Direction}: the expected relationship between the
        mechanism and future returns, such as continuation, reversal, or
        range oscillation;

  \item \textbf{Output}: the form in which the mechanism is expressed as
        a tradable signal, such as a continuous score, event decay, or
        cross-sectional rank.
\end{itemize}

Together, these dimensions specify what market phenomenon is considered,
under which conditions it is interpreted, what evidence supports it, how
it is expected to affect future returns, and how the resulting signal is
expressed. Unlike formula trees or sequential decision processes, a
semantic plan does not impose a predefined ordering or compositional
structure among these dimensions. They jointly define a point in the
trading-semantic space, allowing exploration without introducing
artificial constraints on the underlying market mechanisms.